---
title: "Ubiqum_1_IoTAnalytics_TotalEnergyConsumption_3Submeters"
author: "SuninChoi"
date: "11/15/2019"
output: html_document
Exploratory data analysis on total energy consumption based on 3 sub-meters
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Domain Research and Exploratory Data Analysis

### 1. Obtain the data using SQL query
#### 1.1 open libraries

```{r install RMySQL}
library(RMySQL)
library(ggplot2)
library(funModeling)
library(corrplot)
library(dplyr)
library(readxl)
library(readr)
library(caret)      #R modeling workhorse & ggplot2
library(tidyverse)  #Package for tidying data
library(lubridate)  #For working with dates/times of a time series
library(DEoptimR)
library(gridExtra)
library(car)
library(VIM)        #Visualizing and imputing missing values
library(Hmisc)      #for descriptive statistics
library(forecast)   #forcasting package
library(kableExtra) #fancy table generator
library(broom)      #Tidy statistical summary output
library(knitr)      #report generation
```


#### 1.2 Connect to the database and see what it contains


```{r database and data, echo=FALSE}
# Create a database connection 
con = dbConnect(MySQL(), user='deepAnalytics', password='Sqltask1234!', dbname='dataanalytics2018', host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')

# List the tables contained in the database 
dbListTables(con)

# Use asterisk to specify all attributes for download
irisALL <- dbGetQuery(con, "SELECT * FROM iris")

yr_2006All <- dbGetQuery(con, "SELECT * FROM yr_2006")
yr_2007All <- dbGetQuery(con, "SELECT * FROM yr_2007")
yr_2008All <- dbGetQuery(con, "SELECT * FROM yr_2008")
yr_2009All <- dbGetQuery(con, "SELECT * FROM yr_2009")
yr_2010All <- dbGetQuery(con, "SELECT * FROM yr_2010")
summary(yr_2010All)
df_status(yr_2010All)


# Combine tables into one dataframe using dplyr
yr3 <- bind_rows(yr_2007All, yr_2008All, yr_2009All)
summary(yr3)
head(yr3)
```


### 2. Preprocessing
#### 2.1. Create new columns for date and time attributes


```{r}
## Combine Date and Time attribute values in a new attribute column
yr3 <-cbind(yr3,paste(yr3$Date,yr3$Time), stringsAsFactors=FALSE)

## Give the new attribute in the 6th column a header name 
##!!! NOTE: if you downloaded more than 5 attributes you will need to change the column number)
colnames(yr3)[11] <-"DateTime"

## Move the DateTime attribute within the dataset
yr3 <- yr3[,c(ncol(yr3), 1:(ncol(yr3)-1))]
head(yr3)
```


#### 2.2 Convert the new DateTime attribute to a DateTime data type called POSIXct


```{r}
# Convert DateTime from POSIXlt to POSIXct 
yr3$DateTime <- as.POSIXct(yr3$DateTime, "%Y/%m/%d %H:%M:%S")

# Add the time zone
attr(yr3$DateTime, "tzone") <- "Europe/Paris"

# Inspect the data types
str(yr3)


# Exclude 2010 year data
yr3_1 <- filter(yr3, year(DateTime) != 2010)
```


#### 2.3 Rename variables and assess Values


```{r}
# Exclude 2010 year data
yr3_1 <- filter(yr3, year(DateTime) != 2010)

# Change feature names
colnames(yr3_1)

colnames(yr3_1)[5] <- 'Global_active_power'
colnames(yr3_1)[6] <- 'Global_reactive_power'
colnames(yr3_1)[9] <- 'Kitchen'
colnames(yr3_1)[10] <- 'LaundryRoom'
colnames(yr3_1)[11] <- 'HeaterOrAC'


#-Visualize extent and pattern of missing data
aggr(yr3_1, col=c('navyblue','red'),
                  numbers=TRUE, 
                  sortVars=TRUE, 
                  labels=names(yr3_1),
                  cex.axis=.7, 
                  gap=3, 
                  ylab=c("Histogram of missing data","Pattern"), 
                  digits=2)

#-Remove rows with NA's
yr3_1 <- na.omit(yr3_1)
  
#-Check that there are no missing values remaining
sum(is.na(yr3_1))

```


#### 2.4. Create a new column for 3 submeters


```{r}
# Create long form of data set
yr3_tidy <- yr3_1 %>%
  gather(Meter, Watt_hr, `Kitchen`, `LaundryRoom`, `HeaterOrAC`)  

# Convert meter feature to categorical
yr3_tidy$Meter <- factor(yr3_tidy$Meter)

# Select features
yr3_tidy <- yr3_tidy %>% 
  select(DateTime, Global_active_power, Global_reactive_power, Voltage, Global_intensity, Meter, Watt_hr)

# Date attributes with lubridate
yr3_tidy$year <- year(yr3_tidy$DateTime)
yr3_tidy$quarter <- quarter(yr3_tidy$DateTime)
yr3_tidy$month <- month(yr3_tidy$DateTime)
yr3_tidy$week <- week(yr3_tidy$DateTime)
yr3_tidy$weekdays <- weekdays(yr3_tidy$DateTime)
yr3_tidy$day <- day(yr3_tidy$DateTime)
yr3_tidy$hour <- hour(yr3_tidy$DateTime)

#-peak at data 
glimpse(yr3_tidy)
summary(yr3_tidy)
```


## 3. Explore the data
### 3.1  Sub Meters Analysis
#### 3.1.1 Yearly Time Period

```{r}
library(wesanderson)
names(wes_palettes)

# Yearly Time Period
yr3_tidy %>%
  group_by(year, Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),3)) %>%
  ggplot(aes(x=factor(year), y=sum)) +
  labs(x='Year', y='kWh') +
  ggtitle('Yearly Total Energy Consumption') +
  geom_bar(stat='identity', aes(fill = Meter), color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3))


yr3_tidy %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Year', y='Proportion of Energy Usage') +
  ggtitle('Yearly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3))




```


#### 3.1.2 Quarterly Time Period


```{r}
#-Quarterly bar plot
yr3_tidy %>%
  group_by(quarter(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),3)) %>%
  ggplot(aes(x=factor(`quarter(DateTime)`), y=sum)) +
  labs(x='Quarter of the Year', y='kWh') +
  ggtitle('Quarterly Total Energy Consumption') +
  geom_bar(stat='identity', aes(fill = Meter), color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Moonrise1", n = 3))

yr3_tidy %>%
  group_by(quarter(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`quarter(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Quarter of the year', y='Proportion of Energy Usage') +
  ggtitle('Quarterly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Moonrise1", n = 3))

```


#### 3.1.3 Monthly Time Period


```{r}
#-Monthly bar chart
yr3_tidy %>%
  group_by(month, Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(month), y=sum)) +
    labs(x='Month of the Year', y='kWh') +
    ggtitle('Monthly Total Energy Usage') +
    geom_bar(stat='identity', aes(fill = Meter), colour='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 3))

yr3_tidy %>%
  group_by(month(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`month(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Month of the year', y='Proportion of Energy Usage') +
  ggtitle('Monthly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 3))

```


#### 3.1.4 Weekly and Weekdays Time Period


```{r}
#-Week of the year- bar plot
yr3_tidy %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`week(DateTime)`), y=sum)) +
    labs(x='Week of the Year', y='kWh') +
    ggtitle('Weekly Total Energy Usage') +
    theme(axis.text.x = element_text(angle=90)) +
    geom_bar(stat='identity', aes(fill=Meter), colour='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 3))

yr3_tidy %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`week(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Week of the year', y='Proportion of Energy Usage') +
  ggtitle('Weekly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 3))


yr3_tidy$weekdays <- ordered(yr3_tidy$weekdays, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))


yr3_tidy %>%
  group_by(weekdays, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=weekdays, y=sum)) +
           labs(x='Weekdays', y='kWh') +
           ggtitle('Weekdays Total Energy Usage') +
           theme(axis.text.x = element_text(angle=90)) +
           geom_bar(stat='identity', aes(fill=Meter), colour='black') + 
           theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
           theme(text = element_text(size = 10)) +
           scale_fill_manual(values = wes_palette("Rushmore1", n = 3))

totaldataset %>%
  group_by(Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),4)) %>%
  ggplot(aes(x=Meter, y=sum)) +
  labs(x='Meter', y='kWh') +
  ggtitle('Total Energy Consumption') +
  geom_bar(stat='identity', aes(fill = Meter), color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 4))




```


#### 3.1.5 Daily Time Period


```{r}
#-Hour of day bar chart
yr3_tidy %>%
  filter(month(DateTime) == c(3,5,11,26)) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), y=sum)) +
  labs(x='Hour of the Day', y='kWh') +
  ggtitle('Daily Total Energy Usage') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Chevalier1", n = 3))

yr3_tidy %>%
  group_by(month(DateTime) == c(3, 5, 11, 26)) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Hour of the Day', y='Proportion of Energy Usage') +
  ggtitle('Daily Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Chevalier1", n = 3))

```


### 3.2  Remaining Energy Analysis
#### 3.2.1 Processing


```{r}
# Select features
yr3_tidy <- yr3_tidy %>% 
  select(DateTime, Global_active_power, Global_reactive_power, Voltage, Global_intensity, Meter, Watt_hr)


# Create a new column for remaing energy
reE<- yr3_1 %>% 
  mutate(kw_per_min =(Global_active_power*1000)/60) %>% 
  mutate(restE = kw_per_min 
         - Kitchen - LaundryRoom - HeaterOrAC)

sum(reE$Kitchen, reE$LaundryRoom, reE$HeaterOrAC)
sum(reE$remainingE) 

# Create long form of data set
totaldataset <- reE %>%
  gather(Meter, Watt_hr, `Kitchen`, `LaundryRoom`,`HeaterOrAC`, `restE`)
totaldataset <- na.omit(totaldataset)
summary(totaldataset$Watt_hr)

# Convert meter feature to categorical
totaldataset$Meter <- factor(totaldataset$Meter)

# Date attributes with lubridate
totaldataset$year <- year(totaldataset$DateTime)
totaldataset$quarter <- quarter(totaldataset$DateTime)
totaldataset$month <- month(totaldataset$DateTime)
totaldataset$week <- week(totaldataset$DateTime)
totaldataset$weekdays <- weekdays(totaldataset$DateTime)
totaldataset$day <- day(totaldataset$DateTime)
totaldataset$hour <- hour(totaldataset$DateTime)
```


#### 3.2.2 Explore data
* Total Energy Consumption w/remaining energy


```{r}
totaldataset %>%
  group_by(Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),4)) %>%
  ggplot(aes(x=Meter, y=sum)) +
  labs(x='Meter', y='kWh') +
  ggtitle('Total Energy Consumption') +
  geom_bar(stat='identity', aes(fill = Meter), color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 4))

totaldataset %>%
  group_by(Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=Meter, sum, group=Meter,fill=Meter)) +
  labs(x='Meter', y='Proportion of Energy Usage') +
  ggtitle('Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 4))

# Pie Chart from data frame with Appended Sample Sizes
bp <- totaldataset %>%
  group_by(Meter) %>%
  summarise(sum = sum(Watt_hr)) 

bpp <- ggplot(bp, aes(x="", y=sum, fill=Meter))+
  geom_bar(width = 1, stat = "identity") + 
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 4))
bpp

pie <- bpp + coord_polar("y", start=0)
pie
sum(bp$sum)
# eachsum/totalsum
```


* Time based Analysis w/remaining energy


```{r}
totaldataset %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Year', y='Proportion of Energy Usage') +
  ggtitle('Yearly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 4))

totaldataset %>%
  group_by(quarter(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`quarter(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Quarter of the year', y='Proportion of Energy Usage') +
  ggtitle('Quarterly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Moonrise1", n = 4))

totaldataset %>%
  group_by(month(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`month(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Month of the year', y='Proportion of Energy Usage') +
  ggtitle('Monthly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 4))

totaldataset %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`week(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Week of the year', y='Proportion of Energy Usage') +
  ggtitle('Weekly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 4))


totaldataset$weekdays <- ordered(totaldataset$weekdays, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

totaldataset %>%
  group_by(weekdays, Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`weekdays`), sum, group=Meter,fill=Meter)) +
  labs(x='Week of the year', y='Proportion of Energy Usage') +
  ggtitle('Weekly Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Rushmore1", n = 4))

totaldataset %>%
  group_by(month(DateTime) == c(3, 5, 11, 26)) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Hour of the Day', y='Proportion of Energy Usage') +
  ggtitle('Daily Total Energy Consumption - Percentages') +
  geom_bar(stat='identity', position='fill', color='black') +
  theme(panel.border=element_rect(colour='lightgray', fill=NA)) +
  theme(text = element_text(size = 10)) +
  scale_fill_manual(values = wes_palette("Chevalier1", n = 4))
```




