---
title: "2IoTAnalytics_Task2_TimeSeries"
author: "SuninChoi"
date: "11/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Visualize and Analyze Energy Data

### 1.Visualize the data

```{r}
EC<- yr3_1 %>% 
  mutate(kw_per_min =(Global_active_power*1000)/60) %>% 
  mutate(restE = kw_per_min 
         - Kitchen - LaundryRoom - HeaterOrAC)
#1. Granularity
## Plot all of categories
plot(EC$Kitchen)
plot(EC$LaundryRoom)

EC$year <- year(EC$DateTime)
EC$quarter <- quarter(EC$DateTime)
EC$month <- month(EC$DateTime)
EC$week <- week(EC$DateTime)
EC$weekdays <- weekdays(EC$DateTime)
EC$day <- day(EC$DateTime)
EC$hour <- hour(EC$DateTime)
EC$minute <- minute(EC$DateTime)
EC$second <- second(EC$DateTime)
EC$weekday <- weekday(EC$DateTime)

#2.Subsetting and meaningful time periods
## Subset the second week of 2008 - All Observations
houseWeek <- filter(EC, year == 2008 & week == 2)
## Plot subset houseWeek
plot(houseWeek$Kitchen)

library(plotly)
#3.Visualize a single day with Plotly
## Subset the 9th day of January 2008 - All observations
houseDay <- filter(EC, year == 2008 & month == 1 & day == 9)
## Plot sub-meter 1
plot_ly(houseDay, x = ~houseDay$DateTime, y = ~houseDay$Kitchen, type = 'scatter', mode = 'lines')

## Plot 3sub-meters with title, legend and labels - All observations 
plot_ly(houseDay, x = ~houseDay$DateTime, y = ~houseDay$Kitchen, type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay$LaundryRoom, mode = 'lines') %>%
 add_trace(y = ~houseDay$HeaterOrAC, mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))

#4. Reducing Granularity
## Subset the 9th day of January 2008 - 10 Minute frequency
houseDay10 <- filter(EC, year == 2008 & month == 1 & day == 9 & (minute == 0 | minute == 10 | minute == 20 | minute == 30 | minute == 40 | minute == 50))
## Plot sub-meter 1, 2 and 3 with title, legend and labels - 10 Minute frequency
plot_ly(houseDay10, x = ~houseDay10$DateTime, y = ~houseDay10$Kitchen, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay10$LaundryRoom, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseDay10$HeaterOrAC, name = 'Heater or AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))


## Subset the 9th day of January 2008 - 10 Minute frequency
week32 <- filter(EC, year == 2009 & week == 32 & (minute == 0 | minute == 10 | minute == 20 | minute == 30 | minute == 40 | minute == 50))
## Plot sub-meter 1, 2 and 3 with title, legend and labels - 10 Minute frequency
plot_ly(week32, x = ~week32$DateTime, y = ~week32$Kitchen, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~week32$LaundryRoom, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~week32$HeaterOrAC, name = 'Heater or AC', mode = 'lines') %>%
 layout(title = "Power Consumption 32th Week, 2009",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))

```
* What peaks might represent the water heater? How about the AC? Shower, Cooling places
* What could be happening in the laundry room? Lights, Refriger
* How many times during this day are kitchen appliances being used? 2 times 
* does the data from these three sub-meters contain useful information for the homeowner? 


### 2. Prepare to analyze the data

```{r}
#1. Case1
##1.1 Store your data frame(s) as time series
### Subset to one observation per week on Mondays at 8:00pm for 2007, 2008 and 2009
house070809weekly <- filter(EC, weekdays == 'Monday' & hour == 20 & minute == 1)

## Create TS object with SubMeter3
tsSM3_070809weekly <- ts(house070809weekly$HeaterOrAC, frequency=52, start=c(2007,1))

##. Produce time series plots
## Plot sub-meter 3 with autoplot (you may need to install these packages)
library(ggfortify)
autoplot(tsSM3_070809weekly)

## Plot sub-meter 3 with autoplot - add labels, color
autoplot(tsSM3_070809weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Heater or AC")

## Plot sub-meter 3 with plot.ts
plot.ts(tsSM3_070809weekly)

#Case 2
## Subset to one observation per week on Mondays at 1:00am for 2007, 2008 and 2009 
## why no newyear 0 hour 
housenewyear <- filter(EC, month == 1 & day == 1 & hour == 1 & minute == 1)

## Create TS object with SubMeter3
tsSM3_0101weekly <- ts(housenewyear$HeaterOrAC, frequency=1, start=c(2007,1))

#. Produce time series plots
## Plot sub-meter 3 with autoplot (you may need to install these packages)
autoplot(tsSM3_0101weekly)

## Plot sub-meter 3 with autoplot - add labels, color
autoplot(tsSM3_0101weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Heater or AC")

## Plot sub-meter 3 with plot.ts
plot.ts(tsSM3_0101weekly)

#Case 3
## Subset to one observation per week on Mondays at 1:00am for 2007, 2008 and 2009 
## why no newyear 0 hour 
housedec <- filter(EC, month == 12)

## Create TS object with SubMeter3
tsSM3_housedec <- ts(housedec$HeaterOrAC, frequency=1, start=c(2007,12))

#. Produce time series plots
## Plot sub-meter 3 with autoplot (you may need to install these packages)
autoplot(tsSM3_housedec)

## Plot sub-meter 3 with plot.ts
plot.ts(tsSM3_housedec)
```


3. Forecasting a time series


```{r}
library(forecast)

## Apply time series linear regression to the sub-meter 3 ts object and use summary to 
fitSM3 <- tslm(tsSM3_070809weekly ~ trend + season) 
summary(fitSM3)

## Create the forecast for sub-meter 3. Forecast ahead 20 time periods 
forecastfitSM3 <- forecast(fitSM3, h=20)

## Plot the forecast for sub-meter 3. 
plot(forecastfitSM3)

## Create sub-meter 3 forecast with confidence levels 80 and 90
forecastfitSM3c <- forecast(fitSM3, h=20, level=c(80,90))

## Plot sub-meter 3 forecast, limit y and add labels
plot(forecastfitSM3c, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time")

```


4. Decomposing a Seasonal Time Series


```{r}
# Decomposing a Seasonal Time Series
## Decompose Sub-meter 3 into trend, seasonal and remainder
components070809SM3weekly <- decompose(tsSM3_070809weekly)
## Plot decomposed sub-meter 3 
plot(components070809SM3weekly)
## Check summary statistics for decomposed sub-meter 3 
summary(components070809SM3weekly)

```


## 5. Holt-Winters Forecasting


```{r}
#1. Remove Seasonal Components
## Seasonal adjusting sub-meter 3 by subtracting the seasonal component & plot
tsSM3_070809Adjusted <- tsSM3_070809weekly - components070809SM3weekly$seasonal
autoplot(tsSM3_070809Adjusted)
## Test Seasonal Adjustment by running Decompose again. Note the very, very small scale for Seasonal
plot(decompose(tsSM3_070809Adjusted))

#2. HoltWinters Simple Exponential Smoothing & Plot
## Holt Winters Exponential Smoothing & Plot
tsSM3_HW070809 <- HoltWinters(tsSM3_070809Adjusted, beta=FALSE, gamma=FALSE)
plot(tsSM3_HW070809, ylim = c(0, 25))

## HoltWinters forecast & plot
tsSM3_HW070809for <- forecast(tsSM3_HW070809, h=25)
plot(tsSM3_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 3")

## Forecast HoltWinters with diminished confidence levels
tsSM3_HW070809forC <- forecast(tsSM3_HW070809, h=25, level=c(10,25))

## Plot only the forecasted area
plot(tsSM3_HW070809forC, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 3", start(2010))

```


* TSstudio


```{r}
library(TSstudio)
# Ploting time series object
ts_plot(tsSM3_070809weekly, title = "Energy Consumption in Kitchen", Ytitle = "Watt")

# Seasonal plot
ts_seasonal(tsSM3_070809weekly, type = "all")

# Heatmap plot
ts_heatmap(tsSM3_070809weekly)

# ACF and PACF plots
ts_cor(tsSM3_070809weekly, lag.max = 60)

# Lags plot
ts_lags(tsSM3_070809weekly, lags = 1:12)

# Seasonal lags plot
ts_lags(tsSM3_070809weekly, lags = c(12, 24, 36, 48))

```